In-silico molecular enrichment and clearance of the human intracranial space

### 2025-06-04 paper/main.tex
Address overstatement comment.

% What is the big picture problem and why is it important? 
The mechanisms underlying molecular transport within the intracranial
space are fundamental to human brain health and
function. Neurodegenerative diseases such as Alzheimer's, Parkinson's,
and Huntington's disease are all associated with abnormal accumulation
of protein aggregates together with alterations in transport and
clearance characteristics~\cite{rasmussen2018glymphatic,
  harrison2020impaired, eide2023plasma, liu2024glymphatic}. Moreover,
sleep and conversely sleep-deprivation play a definite yet enigmatic
role in modulating molecular enrichment and
clearance~\cite{xie2013sleep, eide2021sleep, eide2022altered,
  miao2024brain, hauglund2025norepinephrine}. In the last decade,
established theories have been challenged by new findings revealing a
greater degree of molecular movement and
exchange~\cite{iliff2012paravascular, ringstad2017glymphatic,
  louveau2017understanding, proulx2021cerebrospinal,
  bohr2022glymphatic}, as well as substantial variability in
enrichment and clearance between individuals and between patient
cohorts~\cite{ringstad2018brain, eide2021direction, eide2021impaired,
  eide2022altered}. These observations provide distinct opportunities
(and challenges) for personalized medicine e.g.~for tailored
intrathecal delivery of chemotherapy~\cite{lohela2022glymphatic} and
for early diagnostics of impaired brain
clearance~\cite{eide2021clinical, van2024human}. In spite of their
importance, our understanding of these mechanisms is incomplete with
open challenges and significant debate -- in part relating to the
translation of knowledge between scales, species, experimental
protocols, clinical cohorts, and individuals.

>>>>

% What is the big picture problem and why is it important? 
The mechanisms underlying molecular transport within the intracranial
space are fundamental to human brain health and
function. Neurodegenerative diseases such as Alzheimer's, Parkinson's,
and Huntington's disease are all associated with abnormal accumulation
of protein aggregates together with alterations in transport and
clearance characteristics~\cite{rasmussen2018glymphatic,
  harrison2020impaired, eide2023plasma, liu2024glymphatic}. Moreover,
sleep and conversely sleep-deprivation play a definite yet enigmatic
role in modulating molecular enrichment and
clearance~\cite{xie2013sleep, eide2021sleep, eide2022altered,
  miao2024brain, hauglund2025norepinephrine}. In the last decade,
established theories have been challenged by new findings revealing a
greater degree of molecular movement and
exchange~\cite{iliff2012paravascular, ringstad2017glymphatic,
  louveau2017understanding, proulx2021cerebrospinal,
  bohr2022glymphatic}, as well as substantial variability in
enrichment and clearance between individuals and between patient
cohorts~\cite{ringstad2018brain, eide2021direction, eide2021impaired,
  eide2022altered}. These observations provide distinct opportunities
(and challenges) for personalized medicine e.g.~for tailored
intrathecal delivery of chemotherapies\cite{lohela2022glymphatic}
such as methodextrate in acute lymphoblastic leukemia
patients\cite{kadan2009comparison}, and for early diagnostics of
impaired brain clearance\cite{eide2021clinical, van2024human}. In
spite of their importance, our understanding of these mechanisms is
incomplete with open challenges and significant debate -- in part
relating to the translation of knowledge between scales, species,
experimental protocols, clinical cohorts, and individuals.
